---

permalink: /
title: "About"
excerpt: "About me"
author_profile: true
redirect_from: 
  - /about/
  - /about.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I’m a bioinformatics scientist specializing in evolutionary genetics and population biology. I have a Ph.D. in biology from the University of North Carolina and I am currently a postdoctoral researcher in the Whitehead lab at UC Davis. My projects combine evolutionary theory with next generation sequencing data to answer questions at the intersection of basic and applied research.  

# Research
All of my research questions are tied together by a curiosity about molecular changes that make populations unique. Pursuing these questions involves analyses of large –omics data sets which has made me passionate about learning, developing, and applying new bioinformatics approaches. Overall, I strive to conduct research that tests basic theory in evolution and ecology, generates findings that are reproducible and publicly available, and informs population management decisions that preserve biodiversity.

## The genetic basis of complex adaptive traits
Evolutionary radiations showcase astounding examples of adaptation and are often characterized by transitions from generalist to specialist phenotypes. For my dissertation at UNC Chapel Hill, I studied a fascinating radiation of Cyprinodon pupfishes to understand how little genetic differentiation between three young species gave rise to drastic divergence in specialized traits. This radiation consists of a dietary generalist species and two derived trophic niche specialists that feed on novel resources. I discovered candidate alleles under selection that likely contribute to adaptive differences in adaptive craniofacial morphology. I also found evidence suggesting that some of these alleles negatively interact in hybrids to cause gene misregulation, which may act as an important barrier to gene flow during early stages of speciation in sympatry.

## From oceans to rivers and lakes
My postdoctoral research in Andrew Whitehead’s lab continues to explore questions related to adaptation and speciation. For one main project, I’m studying a radiation of Fundulus fishes that occupy diverse osmotic environments to understand adaptations that allowed freshwater specialists to evolve from marine-dwelling ancestors. Since this transition occurred several times independently across multiple clades, I’m searching for parallel and divergent gene regulatory mechanisms underlying freshwater adaptations by identifying cross-species differential gene expression in response to osmotic stress.

## The genomics of population collapse and recovery
For another main project, I’m analyzing genomic samples from several populations of Pacific herring to understand the impacts of the Exxon Valdez oil spill on population collapse and lack of recovery in Prince William Sound, Alaska. The herring fishery in Prince William Sound declined by 75% from 1988 to 1993 and has yet to recover 27 years later. I am using spatial and temporal genomic contrasts to discover changes in allele frequencies pre- and post-collapse that could help to explain the lack of population recovery. 


``````{r, echo=FALSE,eval=FALSE}
library(knitr)
my.jekyll.site <- "joemcgirr.github.io"
KnitPost <- function(input, base.url = my.jekyll.site) {
  opts_knit$set(base.url = base.url)
  fig.path <- paste0("images/", sub(".Rmd$", "", basename(input)), "/")
  opts_chunk$set(fig.path = fig.path)
  opts_chunk$set(fig.cap = "center")
  render_jekyll()
  knit(input, envir = parent.frame())
}

setwd("C:/Users/jmcgirr/Documents/GitHub/my_website/_pages/")
KnitPost("C:/Users/jmcgirr/Documents/GitHub/my_website/_pages/about.Rmd")

```
